Parallel Training of Deep Neural Network using Model Averaging and Butterfly Mixing

### 2015-06-17 paper.tex
add some tables

Distributed model averaging using CPUs is proposed in \cite{zhang2014improving}, and a further improvement 
is done using natural gradient \cite{povey2014parallel}. In this approach, separate models are trained on multiple nodes using 
different partitions of data, and model parameters are averaged after each epoch. It is shown that NG-SGD can effectively improve
convergence and ensure a better model trained using the model averaging framework.

@@

\section{Data parallelization and Model Averging}
Stochastic gradient descent is a popular method for DNN training, even though this is a non-convex optimization 
problem. Since the size of training data is large, minibatch based SGD is often used for model training. 
Roughly, the larger the minibatch size, the higher the converge rate. However, a large minibatch requires
more compute and memory usage, so a straight forward idea for parallelization would be distributing the 
gradient computation to different computing nodes. In each step, the gradient is reduced to one computing node,
averaged and then used to update models in each node. This method can compute the gradient accurately, but 
it requires heavy communication between nodes.

However, if we choose to reduce the weight rather than gradient, it is not necessary to communicate that often. There
is straight forward theory that guarantees convergence, but we do observe reasonable results in experiments.

@@

@@

\section{Initialization}
Conjecture: dbn pretraining initialize the neural network such that it lies in a convex-like area

@@

\subsection{Learing Rate Schedule}

\subsection{Switchboard}

scaling factor
\begin{figure}[htb]
  \centering
  \includegraphics[width=0.5\textwidth]{scaling.png}
  \caption{Scaling factor and speedup factor v.s. number of gpus}
  \label{fig:scaling}
\end{figure}

>>>>

Distributed model averaging is used in \cite{zhang2014improving,miao2014distributed}, and a further improvement 
is done using NG-SGD \cite{povey2014parallel}. In this approach, separate models are trained on multiple nodes using 
different partitions of data, and model parameters are averaged after each epoch. It is shown that NG-SGD can effectively improve
convergence and ensure a better model trained using the model averaging framework.

@@

\section{Data parallelization and Model Averging}
SGD is a popular method for DNN training. Even though neural network training objectives are usually non-convex, 
mini-batch SGD has been shown to be effective for optimizing the objective\cite{seide2011conversational}. 
Roughly speaking, a bigger minibatch size gives a better estimate of the gradient, resulting in a better the converge rate. 
Thus, a straight forward idea for parallellization would be distributing the gradient computation to different computing
nodes. In each step, gradients of minibatches on different nodes are reduced to a single node, averaged and then used to 
update models in each node. This method can compute the gradient accurately, but it requires heavy communication between nodes.
Also, it is shown that increasing minibatch size does not always benefit model training\cite{}, especially in early stage 
of model training.

On the other hand, if we choose to average the parameters rather than gradients, it is not necessary to exchange data that often. 
Currently, there is no straight forward theory that guarantees convergence, but we would like to explore a bit why this strategy 
shoule work, just as we observe in the experiments.

First, in the extreme case where model parameters are averaged after each minibatch update, model averaging is equivalent to 
gradient averaging. If we view the difference of model parameters as gradient in a special space.

Second, it is shown that model averging for convex models is guaranteed to converge \cite{mcdonald2010distributed,mcdonald2009efficient}.
It is suggested that unsupervised pretraining guides the learning towards basins of attraction of minima that support better generalization
from the training data set; \cite{erhan2010does}.

@@

MPI implementation

@@

@@

Reduce frequency

\subsection{Learing Rate Schedule}
Why do we need a large learning rate?

\subsection{Initialization Matters}
Compare random init and dbn init, single and paralle
\begin{table}
  \centering
  \begin{tabular}{c|c|c}
    \hline
     Nodes          &  1    & 16 \\
    \hline
    random init     & --    & --  \\
    \hline
    rbm pretraining & --    & --\\
    \hline
  \end{tabular}
  \caption{Comparing rbm pretraining with random initialization in model averaging setup }
  \label{tab:init}
\end{table}

\subsection{Online NG-SGD Matters}
\begin{table}
  \centering
  \begin{tabular}{c|c|c}
    \hline
     Nodes  &  1    & 16 \\
    \hline
    SGD     & --    & --  \\
    \hline
    NG-SGD  & --    & --\\
    \hline
  \end{tabular}
  \caption{Comparing NG-SGD and naive SGD}
  \label{tab:ngsgd}
\end{table}

\subsection{Speedup Comparison}
scaling factor
\begin{figure}[htb]
  \centering
  \includegraphics[width=0.5\textwidth]{scaling.png}
  \caption{Scaling factor and speedup factor v.s. number of gpus}
  \label{fig:scaling}
\end{figure}
